Wigglyhedra

### 2024-07-04 draft/wigglyhedra.tex
minor changes

\begin{conjecture}
\label{conj:Hamiltonian}
The cover graph of the wiggly lattice~$\wigglyLattice_n$ is Hamiltonian.
\end{conjecture}

Note that unfortunately, wiggly permutations do not form a zigzag language in the sense of~\cite{HartungHoangMutzeWilliams} (which would have guarantied the existence of a Gray code).
\vincent{Ask Aaron what he prefers here...}

@@

\begin{figure}
\centerline{
	\begin{tabular}{c@{\qquad}c@{\qquad}c}
		Spherical objects in bounded 
		&
		&
		\\
		homotopy category of graded projective 
		&
		$\smash{\xrightarrow[\quad\text{combinatorial model}\quad]{}}$
		&
		Curves
		\\
		modules over the zigzag algebra
		&
		&
                \\
          \begin{tikzcd}[ampersand replacement=\&, row sep=-0.1cm, column sep=0.5cm]
                                  \&           \& P_3 \langle -1 \rangle\arrow{dr} \& \\
                                  \&           \&                      \& P_2\\
            P_1\langle -5 \rangle\arrow{r}    \& P_1\langle -3 \rangle\arrow{r}   \& P_1 \langle -1 \rangle\arrow{ur} \& 
          \end{tikzcd}
		&
		&
		\includegraphics[scale=1.3]{curve1}
		\\[.5cm]
		{\scriptsize cohomology} $\Bigg\downarrow$ {\scriptsize filtration}
		&
		&
		$\Bigg\downarrow$ {\scriptsize cut}
		\\[.5cm]
		Heart of the standard $t$-structure
		&
		$\xrightarrow[\quad\text{combinatorial model}\quad]{}$
		&
		Wiggly arcs
		\\
          \begin{tikzcd}[ampersand replacement=\&, row sep=-0.1cm, column sep=0.5cm]
            P_1\langle -1 \rangle \arrow{dr}\& \\
            \& P_2 \\
            P_3 \langle -1 \rangle \arrow{ur}
          \end{tikzcd}
		&
          &
		\includegraphics[scale=1.3]{curve2}
          \\
          \(P_1\)

		&
		&
		\includegraphics[scale=1.3]{curve3}		
	\end{tabular}
}
\caption{An illustration of the Khovanov--Seidel curve model and \cref{prop:cohomology-pieces}.}
\label{fig:curveModel}
\end{figure}

@@

In this section, we give a categorical criterion for two wiggly arcs to be compatible as in~\cref{def:compatible}.
As such, this section is somewhat dual to the previous section --- the criterion in this section determines whether a pair of wiggly arcs can belong together in a single wiggly pseudotriangulation, whereas the previous section generated a wiggly pseudotriangulation given any object of the category.
\begin{proposition}\label{prop:ext1}
  Consider two wiggly arcs \(\alpha\) and \(\alpha'\), and let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible in the sense of~\cref{def:compatible} if and only if \(\Ext^1(X, X') = 0\).
  More generally, the incompatibility degree \(\delta(\alpha,\alpha')\) as in~\cref{def:incompatibilityDegree} is the dimension of \(\Ext^1(X,X')\).
\end{proposition}
\begin{proof}
  The proof follows from a specialisation of the bigraded intersection number calculations that appear in Sections 3 and 4 of~\cite{kho.sei:02}.
  Let \(X\) and \(X'\) be objects in the standard heart that correspond to wiggly arcs \(\alpha\) and \(\alpha'\).
  The bigraded intersection number as defined in~\cite[Sect.~3d]{kho.sei:02} between the curves \(\alpha\) and \(\alpha'\) is a polynomial in the variables \(q_1^{\pm}\), \(q_2^{\pm}\).
  It follows from~\cite[Prop.~4.9]{kho.sei:02} that this bigraded intersection number equals the Poincar{\'e} polynomial of the morphisms between \(X\) and \(X'\), namely the expression
  \begin{equation}
    \label{eq:bigraded-homs}
  \sum_{s_1,s_2}\dim(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1]))q_1^{s_1}q_2^{s_2}.    
\end{equation}
  Note that a morphism in \(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1])\) has level \(s_1 - s_2\); that is, it takes the object \(X\) in level \(0\) to the object \(X'\langle -s_2 \rangle[s_1]\) in level \(s_1 - s_2\).
  We can obtain the dimension of all morphisms of level \(\ell\) by setting \(q_1 = q\) and \(q_2 = q^{-1}\) and taking the coefficient of \(q^{\ell}\) in the expression~\cref{eq:bigraded-homs}.
    The dimension of \(\Ext^1(X,X')\), is therefore the coefficient of \(q\) under this specialisation.

@@

Let \(X\) and \(X'\) be spherical objects in the standard heart that correspond to \(\alpha\) and \(\alpha'\) respectively.
First, note by~\cref{prop:ext1} that \(X\) and \(X'\) are exchangeable if and only if there is a unique non-trivial extension between the corresponding objects of the standard heart.
Next, let us understand \(\beta\) and \(\beta'\) categorically.
\begin{proposition}\label{prop:extensions}
  Let \(\alpha\), \(\alpha'\), \(\beta\), \(\beta'\), and \(X\), \(X'\) be as just discussed.
  Let \(Y\) and \(Y'\) be the objects in the standard heart corresponding to \(\beta\) and \(\beta'\) respectively.
\begin{itemize}
\item if \(\alpha\) and \(\alpha'\) are not pointed, then we have extensions
  \[0 \to X' \to Y \to X \to 0 \quad \text{and}\quad 0 \to X \to Y' \to X' \to 0.\]
\item if \(\alpha\) and \(\alpha'\) are crossing, then we have an extension
  \[0 \to X' \to (Y \oplus Y') \to X \to 0.\]
\end{itemize}
\end{proposition}
\begin{proof}
  The proof is once again a consequence of analysing the extensions explicitly via the complexes from the Khovanov--Seidel reconstruction.
  We do not reproduce the proof here, as the argument is contained in the proof of Theorem 5.5 in~\cite{tho:06}.
  (See also~\cite[Figure 4]{tho:06}.)
\end{proof}
Note that if we write down the relations in the Grothendieck group of our category that correspond to the extensions from each case of~\cref{prop:extensions}, then we obtain precisely the linear dependences that appear between the \(g\)-vectors in~\cref{lem:linearDependences}.
More precisely, we have that
\begin{itemize}
\item if \(\alpha\) and \(\alpha'\) are not pointed, then
  \([Y] = [X] + [X'] = [Y']\),
  from which we see \([X] + [X'] = ([Y] + [Y'])/2\);
\item if \(\alpha\) and \(\alpha'\) are crossing, then
  \([X] + [X'] = [Y] + [Y'].\)
\end{itemize}
This suggests that there should be a way to construct the \(g\)-vectors categorically in such a way that the \(g\)-vectors also map to the Grothendieck group of the category.
At the moment, we do not have a concise construction of the \(g\)-vectors purely categorically, and it would be interesting to find this.

>>>>

\begin{conjecture}
\label{conj:Hamiltonian}
The cover graph of the wiggly lattice~$\wigglyLattice_n$ admits an Hamiltonian path (or even cycle).
\end{conjecture}

Note that unfortunately, wiggly permutations do not form a zigzag language in the sense of~\cite{HartungHoangMutzeWilliams} (which would have guarantied the existence of a Gray code).

@@

\afterpage{
\begin{figure}
\centerline{
	\begin{tabular}{c@{\qquad}c@{\qquad}c}
		Spherical objects in bounded 
		&
		&
		\\
		homotopy category of graded projective 
		&
		$\smash{\xrightarrow[\quad\text{combinatorial model}\quad]{}}$
		&
		Curves
		\\
		modules over the zigzag algebra
		&
		&
                \\
		\begin{tikzcd}[ampersand replacement=\&, row sep=-0.1cm, column sep=0.5cm]
			\&           \& P_3 \langle -1 \rangle\arrow{dr} \& \\
			\&           \&                      \& P_2\\
			P_1\langle -5 \rangle\arrow{r}    \& P_1\langle -3 \rangle\arrow{r}   \& P_1 \langle -1 \rangle\arrow{ur} \& 
		\end{tikzcd}
		&
		&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve1}}
		\\[.5cm]
		{\scriptsize cohomology} $\Bigg\downarrow$ {\scriptsize filtration}
		&
		&
		$\Bigg\downarrow$ {\scriptsize cut}
		\\[.5cm]
		Heart of the standard $t$-structure
		&
		$\xrightarrow[\quad\text{combinatorial model}\quad]{}$
		&
		Wiggly arcs
		\\
		\begin{tikzcd}[ampersand replacement=\&, row sep=-0.1cm, column sep=0.5cm]
			P_1\langle -1 \rangle \arrow{dr}\& \\
			\& P_2 \\
			P_3 \langle -1 \rangle \arrow{ur}
		\end{tikzcd}
		&
          	&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve2}}
		\\
		\(P_1\)
		&
		&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve3}	}
	\end{tabular}
}
\caption{An illustration of the Khovanov--Seidel curve model and \cref{prop:cohomology-pieces}.}
\label{fig:curveModel}
\end{figure}
}

@@

In this section, we give a categorical criterion for two wiggly arcs to be compatible as in~\cref{def:compatible}.
As such, this section is somewhat dual to the previous section --- the criterion in this section determines whether a pair of wiggly arcs can belong together in a single wiggly pseudotriangulation, whereas the previous section generated a wiggly pseudotriangulation given any object of the category.
\begin{proposition}\label{prop:ext1}
  Consider two wiggly arcs \(\alpha\) and \(\alpha'\), and let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to \(\alpha\) and \(\alpha'\) respectively.
  Then \(\alpha\) and \(\alpha'\) are compatible in the sense of~\cref{def:compatible} if and only if \(\Ext^1(X, X') = 0\).
  More generally, the incompatibility degree \(\delta(\alpha,\alpha')\) as in~\cref{def:incompatibilityDegree} is the dimension of \(\Ext^1(X,X')\).
\end{proposition}
\begin{proof}
  The proof follows from a specialisation of the bigraded intersection number calculations that appear in Sections 3 and 4 of~\cite{kho.sei:02}.
  Let \(X\) and \(X'\) be objects in the standard heart that correspond to wiggly arcs \(\alpha\) and \(\alpha'\).
  The bigraded intersection number as defined in~\cite[Sect.~3d]{kho.sei:02} between the curves \(\alpha\) and \(\alpha'\) is a polynomial in the variables \(q_1^{\pm}\), \(q_2^{\pm}\).
  It follows from~\cite[Prop.~4.9]{kho.sei:02} that this bigraded intersection number equals the Poincar{\'e} polynomial of the morphisms between \(X\) and \(X'\), namely the expression
  \begin{equation}
    \label{eq:bigraded-homs}
  \sum_{s_1,s_2}\dim(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1]))q_1^{s_1}q_2^{s_2}.    
\end{equation}
  Note that a morphism in \(\Hom_{\mathcal{C}_m}(X,X'\langle -s_2 \rangle[s_1])\) has level \(s_1 - s_2\); that is, it takes the object \(X\) in level \(0\) to the object \(X'\langle -s_2 \rangle[s_1]\) in level \(s_1 - s_2\).
  We can obtain the dimension of all morphisms of level \(\ell\) by setting \(q_1 = q\) and \(q_2 = q^{-1}\) and taking the coefficient of \(q^{\ell}\) in the expression~\ref{eq:bigraded-homs}.
    The dimension of \(\Ext^1(X,X')\), is therefore the coefficient of \(q\) under this specialisation.

@@

Let \(X\) and \(X'\) be spherical objects in the standard heart that correspond to \(\alpha\) and \(\alpha'\) respectively.
First, note by~\cref{prop:ext1} that \(X\) and \(X'\) are exchangeable if and only if there is a unique non-trivial extension between the corresponding objects of the standard heart.
Next, let us understand~\(\beta\) and \(\beta'\) categorically.
\begin{proposition}\label{prop:extensions}
  Let \(\alpha\), \(\alpha'\), \(\beta\), \(\beta'\), and \(X\), \(X'\) be as just discussed.
  Let \(Y\) and \(Y'\) be the objects in the standard heart corresponding to \(\beta\) and \(\beta'\) respectively.
\begin{itemize}
\item if \(\alpha\) and \(\alpha'\) are not pointed, then we have extensions
  \[0 \to X' \to Y \to X \to 0 \quad \text{and}\quad 0 \to X \to Y' \to X' \to 0.\]
\item if \(\alpha\) and \(\alpha'\) are crossing, then we have an extension
  \[0 \to X' \to (Y \oplus Y') \to X \to 0.\]
\end{itemize}
\end{proposition}
\begin{proof}
  The proof is once again a consequence of analysing the extensions explicitly via the complexes from the Khovanov--Seidel reconstruction.
  We do not reproduce the proof here, as the argument is contained in the proof of Theorem 5.5 in~\cite{tho:06}.
  (See also~\cite[Figure 4]{tho:06}.)
\end{proof}
Note that if we write down the relations in the Grothendieck group of our category that correspond to the extensions from each case of~\cref{prop:extensions}, then we obtain precisely the linear dependences that appear between the \(g\)-vectors in~\cref{lem:linearDependences}.
More precisely, we have that
\begin{itemize}
\item if \(\alpha\) and \(\alpha'\) are not pointed, then
  \([Y] = [X] + [X'] = [Y']\),
  from which we see \([X] + [X'] = ([Y] + [Y'])/2\);
\item if \(\alpha\) and \(\alpha'\) are crossing, then
  \([X] + [X'] = [Y] + [Y'].\)
\end{itemize}
This suggests that there should be a way to construct the \(g\)-vectors categorically in such a way that the \(g\)-vectors also map to the Grothendieck group of the category.
At the moment, we do not have a concise construction of the \(g\)-vectors purely categorically, and it would be interesting to find this.
